IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-10-18 paper/intro.tex

Information flow control (IFC) is a promising technique
that provides security
by tracking the flow of sensitive data through a system.
Untrusted code is confined so that it cannot exfiltrate data, except as
per an information flow policy.  Significant research has been devoted to
adding various forms of IFC to different kinds of programming languages
and systems.  In the context of the web, however, there is a strong
motivation to preserve JavaScript's semantics \Red{and avoid any modifications
  to its engine}, while retrofitting it with dynamic information
flow control.

\Red{\sout{One approach to information flow control is fine-grained tracking,
which offers great flexibility.
However, these systems
require deep integration with the programming language and modify
the semantics in non-trivial ways.  This makes fine-grained systems
less suitable in contexts where preserving an existing language
is important.
%
Moreover, the task of adding fine-grained IFC even to a simple language is
non-trivial~\cite{hritcu2013testing}; when considering a real-world language,
such as JavaScript, this challenge becomes almost insurmountable.}}

%In the alternative \textit{coarse-grained} approach to IFC,
\Red{Operating systems research (e.g.~\cite{Zeldovich:2006}) proposes
\textit{coarse-grained} approaches to IFC, where}
a system is divided into relatively coarse computational units,
each with a single label dictating its security policy.
Only
communication between isolated computational units must be tracked.
%A recent system named COWL~\cite{swapi} adds IFC to the
%browser by identifying \emph{web workers} as an existing computational
%unit.  Communication between workers is mediated using
%IFC to ensure non-interference, preventing sensitive workers from
%influencing the computation of less sensitive workers.
This coarse-grained approach provides a number of advantages:
(1) minimal changes to an
existing programming language when adding IFC, which (2) allows
the reuse of existing programs.  Furthermore, (3)
reduced runtime overhead because checks need only
be performed at isolation boundaries
\Red{instead of at every instruction in the code~(e.g. \cite{JSFlow})}.
%, compared to fine-grained
%systems were the runtime overhead can be impractical~\cite{JSFlow}.
Finally, (4) associating
a single security label with the entire computational unit simplifies
understanding and reasoning about \Red{the security guarantees} of the
system, without \sout{information flow} reasoning about \Red{most of} the
technical details of the semantics of the \Red{underlying} programming language.

In this paper, \Red{we present a framework which translates coarse-grained IFC
ideas into a language-based setting}:
%take coarse-grained labeling~\cite{Zeldovich:2006,
%lio} to its logical extreme:
an information flow control system should
be thought of as multiple instances of completely isolated language
runtimes or \emph{tasks}, with information flow control applied to
inter-task communication.  We describe a formal system in which an IFC
system can be designed once and then applied to any programming language
which has control over external effects (e.g., JavaScript or C with
access to hardware privilege separation).  We formalize this system
using an approach by
Matthews-Findler~\cite{Matthews:2007:OSM:1190216.1190220} for combining
operational semantics and prove non-interference guarantees
that are independent of the choice of specific target language.

@@

Our contributions can be summarized as follows:
\begin{itemize}
  \item We give formal semantics for a core coarse-grained
  dynamic information flow control language free of non-IFC constructs.
  We then show how a large class of target languages can be combined
  with this IFC language and prove that the result provides
  non-interference. (Sections~\ref{sec:retrofit} and \ref{sec:formal})
  \item We provide a proof technique to show the non-interference
  of a concrete semantics for a potentially optimized IFC language
  by means of an isomorphism and show a class of restrictions on
  the IFC language that preserves non-interference. (Sections~\ref{sec:concrete})
  \item We have implemented an IFC system based on these semantics
  for Node.js, and we connect our formalism to another implementation
  based on this work
  for client-side JavaScript in Gecko~\cite{swapi}.
  Furthermore, we outline an implementation for the C programming
  language.
  (Section~\ref{sec:real})
%  \item We show how to extend the Matthews-Findler method from just
%  languages that are not simply expressions evaluating to values,
%  but may be collections of threads executing nondeterministically.
  \item We describe ways to enrich the core IFC language with
  more advanced concepts such as labeled values, labeled references, or a
  notion of clearance. (Section~\ref{sec:extensions})
\end{itemize}

>>>>

Information flow control (IFC) is a promising technique
that provides security
by tracking the flow of sensitive data through a system.
Untrusted code is confined so that it cannot exfiltrate data, except as
per an information flow policy.  Significant research has been devoted to
adding various forms of IFC to different kinds of programming languages
and systems.  In the context of the web, however, there is a strong
motivation to preserve JavaScript's semantics and avoid
JavaScript-engine modifications, while retrofitting it with dynamic information
flow control.

%In the alternative \textit{coarse-grained} approach to IFC,
The Operating Systems community has tackled this challenge (e.g.,
in~\cite{Zeldovich:2006}) by taking a \textit{coarse-grained} approach
to IFC: dividing an application into coarse computational units,
each with a single label dictating its security policy, and only
monitoring communication between them.
%Only communication between isolated computational units must be tracked.
%A recent system named COWL~\cite{swapi} adds IFC to the
%browser by identifying \emph{web workers} as an existing computational
%unit.  Communication between workers is mediated using
%IFC to ensure non-interference, preventing sensitive workers from
%influencing the computation of less sensitive workers.
This coarse-grained approach provides a number of advantages when
compared to the fine-grained approaches typically employed by language-based systems.
First, adding IFC does not require intrusive changes to an
existing programming language, thereby also allowing
the reuse of existing programs.  Second, it has a small
runtime overhead because checks need only
be performed at isolation boundaries
instead of (almost) every program instruction~(e.g.,~\cite{JSFlow}).
%, compared to fine-grained
%systems were the runtime overhead can be impractical~\cite{JSFlow}.
Finally, by associating
a single security label with the entire computational unit simplifies
understanding and reasoning about the security guarantees of the
system, without reasoning about most of the
technical details of the semantics of the underlying programming language.
\Red{ds: this last senstence needs work}

In this paper, we present a framework which brings coarse-grained IFC
ideas into a language-based setting:
%take coarse-grained labeling~\cite{Zeldovich:2006,
%lio} to its logical extreme:
an information flow control system should
be thought of as multiple instances of completely isolated language
runtimes or \emph{tasks}, with information flow control applied to
inter-task communication.  We describe a formal system in which an IFC
system can be designed once and then applied to any programming language
which has control over external effects (e.g., JavaScript or C with
access to hardware privilege separation).  We formalize this system
using an approach by
Matthews-Findler~\cite{Matthews:2007:OSM:1190216.1190220} for combining
operational semantics and prove non-interference guarantees
that are independent of the choice of specific target language.

@@

Our contributions can be summarized as follows: \Red{fixme:}
\begin{itemize}
  \item We give formal semantics for a core coarse-grained
  dynamic information flow control language free of non-IFC constructs.
  We then show how a large class of target languages can be combined
  with this IFC language and prove that the result provides
  non-interference. (Sections~\ref{sec:retrofit} and \ref{sec:formal})
  \item We provide a proof technique to show the non-interference
  of a concrete semantics for a potentially optimized IFC language
  by means of an isomorphism and show a class of restrictions on
  the IFC language that preserves non-interference. (Sections~\ref{sec:concrete})
  \item We have implemented an IFC system based on these semantics
  for Node.js, and we connect our formalism to another implementation
  based on this work
  for client-side JavaScript in~\cite{swapi}.
  Furthermore, we outline an implementation for the C programming
  language.
  (Section~\ref{sec:real})
%  \item We show how to extend the Matthews-Findler method from just
%  languages that are not simply expressions evaluating to values,
%  but may be collections of threads executing nondeterministically.
  \item We describe ways to enrich the core IFC language with
  more advanced concepts such as labeled values, labeled references, or a
  notion of clearance. (Section~\ref{sec:extensions})
\end{itemize}
